---
pdf_document: default
author: "Team 2 Group 11"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
title: "Bivariate statistical analysis of relevant variables"
header-includes:
- \usepackage{lscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing of necessary packages:

```{r, warning=FALSE}
if(!require(plotly))install.packages("plotly"); require(plotly)
```

Loading dataset:

```{r}
mydata = read.csv("../../primera_entrega/datasets/preprocessed_dataset.csv")

```

#### Succes Rate / attacktype1_txt

First we want to analyze if there is a dependence between the variables attacktype1_txt and success.
For that we make a table with values of these two variables.

```{r}
aux = table(mydata$attacktype1_txt,mydata$success);
SuccessRate=numeric(8);
for (x in 1:8) {
  aux2=aux[x,2]/(aux[x,1]+aux[x,2]);
  SuccessRate[x]=aux2;
}
t <- cbind(aux,SuccessRate);
print(t)
```

And then we do the chi-square test, to check if there is a dependency between this two variables or not.

```{r, warning=FALSE}
chisq.test(aux)
```

With a p-value of 2.2e-16 we can reject this test's null hypothesis and conclude this variables are in fact dependent.

Now we want to analyze with grafical representation the number of successful acts of each type of attack.

```{r}
success_attacks <- mydata[mydata$success == 'TRUE',]

f<-table(success_attacks$attacktype1_txt)
f1=as.data.frame(f)

# Pie Chart con Plotly
library(plotly)
plot_ly(f1, labels = ~Var1, values = ~Freq, type = 'pie') %>%
  layout(title = 'Pie Chart of % success attakcs for each attacktype1_txt',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

```

#### Success Rate / weaptype1_txt

First we want to analyze if there is a dependence between the variables weaptype1_txt and success.
For that we make a table with values of these two variables.

```{r}
aux = table(mydata$weaptype1_txt,mydata$success);

SuccessRate=numeric(11);
for (x in 1:11) {
  aux2=aux[x,2]/(aux[x,1]+aux[x,2]);
  SuccessRate[x]=aux2;
}
t <- cbind(aux,SuccessRate);
print(t)
```

And then we do the chi-square test, to check if there is a dependency between this two variables or not.

```{r, warning=FALSE}
chisq.test(aux)
```


With a p-value of 2.2e-16 we can reject this test's null hypothesis and conclude this variables are in fact dependent.

Now we want to analyze with grafical representation the number of successful acts of each weapon type.

```{r}
success_attacks <- mydata[mydata$success == 'TRUE',]

f<-table(success_attacks$weaptype1_txt)
f1=as.data.frame(f)

# Pie Chart con Plotly
library(plotly)
# HACER QUE NO SE VEA RECORTADO!!! 
plot_ly(f1, labels = ~Var1, values = ~Freq, type = 'pie', height = 680) %>%
  layout(title = 'Pie Chart of % success attakcs for each weaptype1_txt')

```

#### targtype1_txt / provstate

In this section we want see if there is a relation between different type of targets and the state where was an attack.

```{r, warning=FALSE}
aux = table(mydata$provstate,mydata$targtype1_txt);
chisq.test(aux)
```

With a p-value of 2.2e-16 we can reject this test's null hypothesis and conclude this variables are in fact dependent. 

So with the next table we can see that some big states as California, New York, Florida or Illinois and Puerto Rico have more variability in the type of targets.

```{r}
aux2 = table(mydata$targtype1_txt,mydata$provstate);
print(aux2)
```

And with this pie chart we can see that the most popular/common target type are:

Business - individuals or organizations engaged in commercial or mercantile activity as a means of livelihood.

Private Citizens & Property - includes attacks on individuals, the public in general or attacks in public areas including markets, commercial streets, busy intersections and pedestrian malls.

Government (General) - any attack on a government building; government member, former members, including members of political parties in official capacities, their convoys, or events sponsored by political parties; political movements; or a government sponsored institution where the attack is expressly carried out to harm the government.

```{r}
f2=as.data.frame(aux2)
plot_ly(f2, labels = ~Var1, values = ~Freq, type = 'pie') %>%
  layout(title = 'Pie Chart of different targtype1_txt',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```


#### Success Rate / targtype1_txt

Here we want to study if there is a relation between targtype1_txt and success.
For this we make chisq.test() with the next results:

```{r}
aux = table(mydata$targtype1_txt ,mydata$success)
chisq.test(aux)
```

With a p-value of 2.704e-05 we can reject this test's null hypothesis and conclude this variables are in fact dependent. 

And with this table we can see the success rate of each target types implicated in attacks.

```{r}
SuccessRate=numeric(21);
for (x in 1:21) {
  aux2=aux[x,2]/(aux[x,1]+aux[x,2]);
  SuccessRate[x]=aux2;
}
t <- cbind(aux,SuccessRate);
print(t)
```

#### nperpcap / claimmode_txt

We think it might be interesting to know whether the claimed mode affects or helps capturing a perpetrator.
To analyse it we use the next table, which contains claimed mode references and number of perpetrators captured, and the correspondent boxplot chart.

```{r}
aux = table(mydata$claimmode_txt, mydata$nperpcap)
aux = as.data.frame(aux)
mat = matrix(aux[,"Freq"],9,11,FALSE,list(unique(aux[,"Var1"]),unique(aux[,"Var2"])))

print(mat)
```

```{r}
# Don't need column with number of 0 nperpcap
mat <- mat[,-1]

rownames(mat) <- c("C-Post","C-Pre","E-mail","Letter","Note", "Other","Pers.","Posted","Video")

boxplot(mat, main = "Boxplot of nperpcap in funtion of claimmode_txt",
        use.cols = FALSE)
```

Legend:
"C-Post: Call (post-incident)", 
"C-Pre: Call (pre-incident)",
"E-mail: E-mail",
"Letter: Letter",
"Note: Note left at scene",
"Other: Other", 
"Pers.: Personal claim", 
"Posted: Posted to website, blog, etc.", 
"Video: Video"

With this boxplot we can obiously see that "Personal claim" have the biggest number of perpetrators captured. We don't consider it's outlier, but we want to see how will change the boxplot if we delete this particular case. 
Soo in the next boxplot we decide to ignore this concrete claimed mode.

```{r}
mat <- mat[-7,]

boxplot(mat, main = "Boxplot of nperpcap in funtion of claimmode_txt without 'Pers.' mode",
        use.cols = FALSE)
```

Legend:
"C-Post: Call (post-incident)", 
"C-Pre: Call (pre-incident)",
"E-mail: E-mail",
"Letter: Letter",
"Note: Note left at scene",
"Other: Other", 
"Posted: Posted to website, blog, etc.", 
"Video: Video"

And we can see that "Letter" and "Note left at scene" are the next two claimed modes with the biggest number of captured perpetrators.

####  nkill / weaptype1_txt

In this part we study if there is a dependence between weapon type and number of kills.
First of all we make a table with corresponding variables and see how our data looks like.

```{r}
aux = table(mydata$weaptype1_txt, mydata$nkill)
print(aux)
```

With this table we understand that there are some columns that have no sense for the study we want to do, soo we decide remove them and prepare our data for boxplot.

```{r}
remove <- c(44,50,190)
Biologicalkill = mydata$nkill[mydata$weaptype1_txt == 'Biological']
Biologicalkill=Biologicalkill[! Biologicalkill %in% remove]

Chemicalkill = mydata$nkill[mydata$weaptype1_txt == 'Chemical']
Chemicalkill=Chemicalkill[! Chemicalkill %in% remove]

Explosiveskill = mydata$nkill[mydata$weaptype1_txt == 'Explosives']
Explosiveskill=Explosiveskill[! Explosiveskill %in% remove]

FakeWeaponskill = mydata$nkill[mydata$weaptype1_txt == 'Fake Weapons']
FakeWeaponskill=FakeWeaponskill[! FakeWeaponskill %in% remove]

Firearmskill = mydata$nkill[mydata$weaptype1_txt == 'Firearms']
Firearmskill=Firearmskill[! Firearmskill %in% remove]

Incendiarykill = mydata$nkill[mydata$weaptype1_txt == 'Incendiary']
Incendiarykill=Incendiarykill[! Incendiarykill %in% remove]

Meleekill = mydata$nkill[mydata$weaptype1_txt == 'Melee']
Meleekill=Meleekill[! Meleekill %in% remove]

Otherkill = mydata$nkill[mydata$weaptype1_txt == 'Other']
Otherkill=Otherkill[! Otherkill %in% remove]

Radiologicalkill = mydata$nkill[mydata$weaptype1_txt == 'Radiological']
Radiologicalkill=Radiologicalkill[! Radiologicalkill %in% remove]

SabotageEquipmentkill = mydata$nkill[mydata$weaptype1_txt == 'Sabotage Equipment']
SabotageEquipmentkill=SabotageEquipmentkill[! SabotageEquipmentkill %in% remove]

Vehiclekill = mydata$nkill[mydata$weaptype1_txt == 'Vehicle']
Vehiclekill=Vehiclekill[! Vehiclekill %in% remove]


boxplot(Biologicalkill,Explosiveskill,Firearmskill,Incendiarykill,Meleekill, main = "Boxplot of kills by weapon type",names=c("Biological","Explosives","Firearms","Incendiary","Melee"),
        use.cols = TRUE)
```

The vas majority of the cases have 0 kills so the graphical representation does not give us much information.

#### nwound / weaptype1_txt

The same type of study as in "nkill / weaptype1_txt" section we repeat it with "nwound / weaptype1_txt".

```{r}
aux = table(mydata$weaptype1_txt, mydata$nwound)
print(aux)
``` 

As we can see, most of the weapon types have small quantities of kills which would distort our representation. We will remove those types for the graphical comparison.
```{r}
Explosiveswound = mydata$nwound[mydata$weaptype1_txt == 'Explosives']
Firearmswound = mydata$nwound[mydata$weaptype1_txt == 'Firearms']
Incendiarywound = mydata$nwound[mydata$weaptype1_txt == 'Incendiary']
Meleewound = mydata$nwound[mydata$weaptype1_txt == 'Melee']
boxplot(Explosiveswound,Firearmswound,Incendiarywound,Meleewound, main = "Boxplot of wounded people by weapon type",names=c("Explosives","Firearms","Incendiary","Melee"),
        use.cols = TRUE)
```

The vas majority of the cases have 0 wounded so the graphical representation does not give us much information.

#### n ยบ acts / year date  

In this analysis we want to look into the distribution of acts over the years. We will see 10 year intervals and 1970-2017 graphs in order to compare them properly.

```{r}
mydata$date = as.Date(mydata$date, "%Y-%m-%d")
mydata$year = as.numeric(format(mydata$date, "%Y"))
aux = table(mydata$year)
#dev.off()
x11()
Conf2x2 = matrix(c(1:6), nrow=3, byrow=FALSE)
layout(Conf2x2)
barplot(aux[0:47], main = "from 1970 to 2017")
barplot(aux[0:10], main = "from 1970 to 1979")
barplot(aux[11:20], main = "from 1980 to 1989")
barplot(aux[21:29], main = "from 1990 to 1999", sub = "(no data from 1993)")
barplot(aux[30:39], main = "from 2000 to 2009")
barplot(aux[39:47], main = "from 2010 to 2017")
x11()
```

We can see that the year with most acts of terrorism is by far 1970. In the last graph 2009-2017 we can clearly see that there is an increasing tendency.

#### n ยบ acts / month

In this analysis we want to see what is the distribution of acts of terrorism over the months in order to see if there is any anomaly.

```{r}
mydata$date = as.Date(mydata$date, "%Y-%m-%d")
mydata$month = as.numeric(format(mydata$date, "%m"))
aux = table(mydata$month);
barplot(aux, main = "Number of terrorism acts by month", names.arg = c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep', 'Oct', 'Nov', 'Dec'))
```

The acts are fairly uniform distributed, with the preferred month being May and the least favorite being November.

#### provstate / gname

This two categorical variable seems interesting to analyze, but first we need to bee sure if there is a relation between them. As we know  for this we can use chi square test.

```{r, warning=FALSE}
chisq.test(mydata$provstate, mydata$gname)
```

With a p-value of 2.2e-16 we can reject this test's null hypothesis and conclude this variables are in fact dependent.

#### gname / provstate (top 3 provstate)

In this analysis we want to see the distributions of acts of terrorism by each group in the 3 most striked states.

We will remove the gname with less than 9 apperances in order to properly view the result as we will not consider them relevant to our study.

```{r}
aux = mydata$gname[mydata$provstate == 'California']
aux=as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie',height = 500) %>%
  layout(title = 'Pie Chart of different gname in California',legend = list(orientation = "h", xanchor = "center", x = 0.5))
```

```{r}
aux = mydata$gname[mydata$provstate == 'New York']
aux=as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie',height = 500) %>%
  layout(title = 'Pie Chart of different gname in New York',
         legend = list(orientation = "h", xanchor = "center", x = 0.5))
```

```{r}
aux = mydata$gname[mydata$provstate == 'Puerto Rico']
aux=as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie') %>%
  layout(title = 'Pie Chart of different gname in Puerto Rico',
         legend = list(orientation = "h", xanchor = "center", x = 0.5))

```

#### provstate / gname (top 3 gname)

In this analysis we want to identify in which states a given gname has committed acts of terrorism. This will provide us insight of the group's target population.

```{r}
aux = mydata$provstate[mydata$gname == 'Anti-Abortion extremists']
aux=as.data.frame(table(aux))
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie', height = 530)
```

We can see that this group widely distributed theirs acts of terrorism over the states.

```{r}
aux = mydata$provstate[mydata$gname == 'Left-Wing Militants']
aux=as.data.frame(table(aux))
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie')
```

We can see that this group's acts were fairly distributed although half of them were targeted to the two main victim states.

```{r}
aux = mydata$provstate[mydata$gname == 'Fuerzas Armadas de Liberacion Nacional (FALN)']
aux=as.data.frame(table(aux))
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie')
```

In this case we can clearly see that the acts were targeted towards the states of New York, Illinois and Puerto Rico.


#### Event Rate / state_governor_party (top 3 provstates)

We want to know what is the percentatge of acts committed while the rule of each state_governor_party. Only the 3 states with most recorded acts will be considered because in some of the states the sample size would be too small.

We will remove the minor state_governor_parties with less than 9 apperances in order to properly view the result as we will not consider them relevant to our study.

First we will check California as it is the first state with most recorded acts of terrorism (532).

```{r}
aux = mydata$state_governor_party[mydata$provstate == 'California']
aux = as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
aux = aux[aux$Freq < 200,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie')
```

Not considering the outliers, we can see that 69.9% of the acts where committed in this state while the Democrat party was ruling.

We will do the same analysis with New York as it is the second state with most recorded acts of terrorism (473).

```{r}
aux = mydata$state_governor_party[mydata$provstate == 'New York']
aux = as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie')
```

In this case the ratio is more equaly distributed, with 51.6% for the Democrats

Lastly we will do the same analysis with Puerto Rico as it is the third state with most recorded acts of terrorism (226).

```{r}
aux = mydata$state_governor_party[mydata$provstate == 'Puerto Rico']
aux = as.data.frame(table(aux))
aux = aux[aux$Freq > 9,]
plot_ly(aux, labels = ~aux, values = ~Freq, type = 'pie')
```

In the pie we can see that 65.3% of the acts were commited while the New Progressive Party was ruling.

#### Success Rate / president_party

In this analysis we want to check what is the ratio of success of terrorism acts while each president_party was ruling.

```{r}
success = mydata$president_party[mydata$success == TRUE]
fail = mydata$president_party[mydata$success == FALSE]
success = as.data.frame(table(success))
fail = as.data.frame(table(fail))
s = success[2]+fail[2]
ts = 100*success[2]/s
tf = 100*fail[2]/s

s=c(ts[1,1],ts[2,1])
f=c(tf[1,1],tf[2,1])
t <- cbind(s,f);
colnames(t) <- c("Success", "Fail")
rownames(t) <- c("Democratic", "Republican")

print(t)
```

As we can see, both parties have similar success rates.
